\section{Introduction}

Static Application Security Testing (SAST) tools are a cornerstone of secure
software development, yet their practical value is undermined by noise.
The Ghost Security CAST report found a 99.5\% false positive rate for
command-injection findings in Python/Flask applications, with 91\% of all
2{,}166 findings across categories classified as noise~\cite{ghost2025cast}.
The picture is similar at scale: NIST's SATE~V and SATE~VI evaluations
measured false positive rates of 70--92\% depending on language and
tool~\cite{delaitre2018satev,delaitre2023satevi}. On the detection side,
the Fluid Attacks benchmark showed that the best fully automated tool found
only 22.7\% of known vulnerabilities, with an average F-score of 1.9\%
across 36 commercial and open-source scanners~\cite{fluidattacks2025}.
Practitioners, faced with hundreds of alerts they must manually triage,
increasingly ignore scanner output altogether.

A new generation of LLM-powered security scanners promises to change this.
By leveraging large language models for semantic code understanding rather than
pattern matching, these tools claim deeper reasoning about data flow, context,
and exploitability. Yet no open benchmark exists to verify whether LLM-based
scanners actually deliver on these promises---or whether they simply trade one
set of failure modes for another. Existing benchmarks fall short on one or more
critical dimensions:

\begin{itemize}
  \item \textbf{Synthetic code.} The OWASP Benchmark~\cite{owaspbenchmark}
    consists of synthetic Java test cases that do not reflect the structure,
    complexity, or idioms of real-world applications.
  \item \textbf{Closed data.} SastBench~\cite{feiglin2026sastbench} is the
    closest prior work to ours, but its dataset has not been publicly released
    and its evaluation focuses on triage agents rather than scanner accuracy.
  \item \textbf{Vendor self-assessment.} Benchmarks published by
    ZeroPath~\cite{zeropath2024}, Cycode~\cite{cycode2024benchmark}, and
    DryRun~\cite{dryrun2025sast} evaluate their own tools on proprietary
    datasets, making independent reproduction impossible.
  \item \textbf{No false positive measurement.} Most benchmarks report
    detection rates alone, ignoring the false positive burden that dominates
    practitioner experience.
\end{itemize}

\noindent
We expand on these limitations in Section~\ref{sec:related-work}. The gap is
clear: no prior work systematically compares rule-based SAST, general-purpose
LLMs, and security-specialized scanners on identical real-world code, measuring
not just what they find but also what they falsely flag.

This paper makes three contributions:

\begin{enumerate}
  \item \textbf{A systematic comparison of 15 scanners across three categories}
    on identical ground truth: 3 rule-based SAST tools (pattern matching),
    11 general-purpose LLM scanner configurations spanning 7 model families
    (broad reasoning applied to security), and 1 security-specialized system
    (purpose-built vulnerability detection).
  \item \textbf{The \textsc{RealVuln} benchmark} that enables this comparison---the
    first fully open-source SAST evaluation framework built from real-world code,
    comprising 26 intentionally vulnerable Python repositories with 796
    hand-labeled findings and 120 false-positive traps, scored using an
    F3-weighted metric ($\beta{=}3$, recall weighted 9$\times$ over precision)
    as the primary metric.
  \item \textbf{A living benchmark} with versioned scoring, a public
    dashboard, and a clear contribution path so that the community can add new
    scanners, repositories, and ground-truth labels over time.
\end{enumerate}

Our evaluation reveals a clear three-tier performance hierarchy. The security-specialized scanner (Kolega) achieves F3\,=\,73.0 (strict), outperforming the best general-purpose LLM (Claude Sonnet~4.6, F3\,=\,51.7) by 41\%, which in turn outperforms the best rule-based SAST tool (Semgrep, F3\,=\,17.7) by nearly 3$\times$. These results suggest that applying general-purpose AI to security prompts, while a major advance over pattern matching, is not sufficient --- purpose-built security architecture delivers meaningfully higher recall.

\paragraph{Disclosure.} One of the evaluated scanners (Kolega) is developed by the authors of this paper. Kolega achieves the highest F3 score in our evaluation, which heightens the importance of independent verification. To mitigate bias, we open-source all ground-truth labels, scoring scripts, raw scanner outputs, and the evaluation harness --- enabling full independent reproduction. The ground-truth labels were created following a systematic labeling protocol applied uniformly across all repositories, without reference to any scanner's output. We actively encourage third-party audits of our methodology and results.

We release all code, ground-truth data, scanner results, scoring tooling, and
the interactive dashboard at
\url{https://github.com/kolega-ai/RealVulnBenchmark}.
