\section{Background \& Related Work}
\label{sec:related-work}

Table~\ref{tab:benchmark-comparison} summarizes how RealVuln relates to prior
benchmarks across seven desirable properties. We discuss each category of prior
work below, highlighting the specific gaps that motivated our design.

%% ── Comparison table ─────────────────────────────────────────────────────────
\begin{table*}[t]
\caption{Comparison of SAST benchmarks. \emph{Real Code}: uses unmodified
  application source, not synthetic snippets. \emph{Open Data}: ground truth
  and scanner results publicly available. \emph{Open Scoring}: scoring code
  released for reproduction. \emph{FP Testing}: explicitly measures false
  positives. \emph{Multi-Scanner}: evaluates $\geq$3 scanners head-to-head.
  \emph{LLM Scanners}: includes LLM-based tools. \emph{Multi-Lang}: covers
  more than one programming language.}
\label{tab:benchmark-comparison}
\centering
\footnotesize
\begin{tabular}{@{}lccccccc@{}}
\toprule
\textbf{Benchmark} & \textbf{Real} & \textbf{Open} & \textbf{Open} & \textbf{FP} & \textbf{Multi-} & \textbf{LLM} & \textbf{Multi-} \\
 & \textbf{Code} & \textbf{Data} & \textbf{Scoring} & \textbf{Test} & \textbf{Scanner} & \textbf{Scan.} & \textbf{Lang} \\
\midrule
OWASP Benchmark~\cite{owaspbenchmark}
  & $\times$ & \checkmark & \checkmark & \checkmark & \checkmark & $\times$ & $\times$ \\
Juliet Test Suite~\cite{boland2012juliet}
  & $\times$ & \checkmark & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\
SastBench~\cite{feiglin2026sastbench}
  & \checkmark & $\times$ & $\times$ & \checkmark & $\times$ & $\times$ & \checkmark \\
ZeroPath/XBOW~\cite{zeropath2024}
  & $\times$ & $\times$ & $\times$ & \checkmark & \checkmark & \checkmark & $\times$ \\
Fluid Attacks~\cite{fluidattacks2025}
  & \checkmark & $\times$ & $\times$ & \checkmark & \checkmark & $\times$ & $\times$ \\
CASTLE~\cite{dubniczky2025castle}
  & $\times$ & \checkmark & \checkmark & $\times$ & \checkmark & \checkmark & \checkmark \\
Cycode~\cite{cycode2024benchmark}
  & \checkmark & $\times$ & $\times$ & $\times$ & \checkmark & $\times$ & \checkmark \\
DryRun~\cite{dryrun2025sast}
  & \checkmark & $\times$ & $\times$ & $\times$ & \checkmark & $\times$ & \checkmark \\
\midrule
\textbf{RealVuln (ours)}
  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & $\times$ \\
\bottomrule
\end{tabular}
\end{table*}

%% ── §1  Synthetic benchmarks ────────────────────────────────────────────────
\paragraph{Synthetic benchmarks.}
The OWASP Benchmark~\cite{owaspbenchmark} is the most widely cited SAST
evaluation suite. It consists of approximately 2{,}740 synthetic Java test
cases, each containing exactly one vulnerability or one false positive marker,
scored via Youden's J statistic (TPR $-$ FPR). While its inclusion of FP
testing was ahead of its time, the synthetic nature of the code---each test
case is a short, self-contained servlet---means that pattern-matching tools
can achieve high scores without understanding real application semantics such
as cross-file data flow, framework routing, or ORM queries. The NIST Juliet
Test Suite~\cite{boland2012juliet} follows the same philosophy for C/C++ and
Java: thousands of small programs, each exercising a single CWE, with ``good''
and ``bad'' variants. Its narrow CWE coverage (focused on buffer errors and
injection) and lack of multi-file context further limit ecological validity.
Both suites remain useful for regression testing individual rules but are
insufficient for evaluating whole-application analysis capabilities.

%% ── §2  Unreleased datasets ─────────────────────────────────────────────────
\paragraph{Unreleased datasets.}
SastBench~\cite{feiglin2026sastbench} is the closest prior work to RealVuln in
its methodology. It assembles 2{,}737 ground-truth samples across 38
programming languages from real CVEs and uses a SAST proxy to generate
realistic false positive candidates, scored with the Matthews Correlation
Coefficient (MCC). Its scope is impressive, and the MCC metric is a principled
choice for imbalanced classification. However, the dataset has not been
publicly released, and its evaluation targets agentic \emph{triage} workflows
rather than raw scanner accuracy: the benchmark measures how well an LLM agent
can classify SAST alerts as true or false positives, rather than whether the
underlying scanner detects vulnerabilities in the first place. This makes
SastBench complementary to RealVuln---it evaluates the post-scan triage stage
that we deliberately exclude---but it cannot serve as an open benchmark for
scanner detection performance.

%% ── §3  Vendor benchmarks ───────────────────────────────────────────────────
\paragraph{Vendor benchmarks.}
Several security vendors have published benchmark results that provide useful
data points but are not independently reproducible. ZeroPath~\cite{zeropath2024}
evaluated its own LLM-based scanner (XBOW) against traditional tools on a
curated set of Python and PHP applications, reporting high detection rates;
however, the ground truth is only partially disclosed and the scoring
methodology is not released, precluding independent verification.
Fluid Attacks~\cite{fluidattacks2025} stands out for its thoroughness: it
tested 36 commercial and open-source scanners against a single large
JavaScript/TypeScript application with 1{,}093 known vulnerabilities found by
human pentesters. The results were sobering---the average F2 score across all
tools was 1.9\%, and the best fully automated scanner achieved only 22.7\%
recall, while the pentesters found 99\%. Unfortunately, neither the application
source nor the ground truth is publicly available. Cycode~\cite{cycode2024benchmark}
and DryRun~\cite{dryrun2025sast} each published multi-scanner evaluations on
real codebases across multiple languages, but neither released the underlying
datasets or scoring code. The fundamental limitation shared by all vendor
benchmarks is a structural conflict of interest: each vendor controls which
applications are tested, which findings are considered ground truth, and how
scoring is performed, making it impossible to rule out selection effects that
favor the vendor's own tool.

%% ── §4  Dataset quality ─────────────────────────────────────────────────────
\paragraph{Dataset quality and label noise.}
A parallel line of research has exposed the fragility of automatically
constructed vulnerability datasets, motivating our choice of hand-labeled
ground truth. Ding et al.~\cite{ding2024primevul} demonstrated that a 7B
parameter language model could achieve 68\% F1 on the widely used BigVul
dataset but only 3\% on their cleaned PrimeVul dataset, attributing the
collapse to pervasive label noise in commit-based labeling (tangled commits,
incomplete fixes, and mislabeled functions). CleanVul~\cite{li2025cleanvul}
applied LLM-based heuristics to filter noisy labels and found that over 40\%
of entries in existing datasets were mislabeled or ambiguous.
CVEFixes~\cite{bhandari2021cvefixes} and ReposVul~\cite{wang2024reposvul}
automated the collection of vulnerability-fixing commits at scale but
inherited the same tangled-commit problem. These findings underscore a key
lesson: \emph{automated label generation at scale trades accuracy for
coverage}. RealVuln takes the opposite position---796 hand-labeled findings
across 26 repositories---accepting lower scale in exchange for label quality
that can be audited line by line.

%% ── §5  Agentic and LLM benchmarks ─────────────────────────────────────────
\paragraph{Agentic and LLM-oriented benchmarks.}
Several recent benchmarks have influenced our design, even where their scope
differs from ours. SWE-Bench~\cite{jimenez2024swebench} introduced the
paradigm of evaluating AI agents on real GitHub issues in Docker-isolated
environments, establishing that agent-agnostic infrastructure and reproducible
evaluation harnesses are both achievable and essential for credibility;
the subsequent SWE-Bench Verified~\cite{openai2025swebenchverified} effort
further demonstrated the value of human-validated labels.
CVE-Bench~\cite{zhu2025cvebench} adapted this approach to offensive security,
measuring whether AI agents can exploit real CVEs in containerized web
applications, and highlighted that model knowledge cutoffs introduce a
confound that benchmarks must control for.
CASTLE~\cite{dubniczky2025castle} is perhaps the closest LLM-focused
benchmark: it evaluates 23 SAST tools and multiple LLMs on CWE detection
across multiple languages. However, CASTLE uses synthetically constructed
micro-benchmarks rather than real applications, placing it in the same
ecological-validity category as OWASP and Juliet.
Yildiz et al.~\cite{yildiz2025jitvul} benchmarked LLM-based agents on
repository-level vulnerability detection and found that agentic approaches did
not yet outperform simple function-level prompting, suggesting that the
bottleneck is not tool orchestration but fundamental model capability---a
finding consistent with our own results (Section~\ref{sec:results}).

%% ── §6  Honest self-critique ────────────────────────────────────────────────
\paragraph{Positioning and limitations of RealVuln.}
Critically, no prior benchmark systematically compares all three scanner
categories---rule-based SAST, general-purpose LLMs, and security-specialized
systems---on identical ground truth. As Table~\ref{tab:benchmark-comparison}
shows, RealVuln is the only benchmark that simultaneously uses real code,
releases all data and scoring code, tests for false positives, and includes
scanners from all three categories.
However, we are forthright about its limitations in v1. First, RealVuln covers
Python only; while this enables depth (Django, Flask, FastAPI, and raw Python
idioms are all represented), it cannot speak to scanner performance on memory-unsafe
languages or compiled ecosystems. Second, our ground truth covers only
OWASP-style application vulnerabilities (injection, XSS, authentication
flaws)---we do not yet include supply-chain, configuration, or
infrastructure-as-code findings. Third, we are ourselves a vendor (Kolega),
which creates the same conflict of interest we criticize above. We mitigate
this by open-sourcing \emph{everything}: ground truth labels, all scanner
result files, the scoring pipeline, and the interactive dashboard, so that any
researcher can audit, challenge, or extend our results. We view transparency
not as a defense against bias but as an invitation for the community to find it.
