\section{Benchmark Design}
\label{sec:benchmark-design}

This section describes the design of \textsc{RealVuln}~v1.0: the taxonomy of target types it recognizes, how the dataset was constructed and labeled, the algorithm used to match scanner findings against ground truth, and the scoring metrics reported.

%% ----------------------------------------------------------------
\subsection{Target Type Taxonomy}
\label{sec:target-types}

We define five \emph{target types} to classify the code under test.
Version~1.0 of \textsc{RealVuln} covers only Type~1; the remaining types are defined here so that the roadmap in Section~7 can reference them precisely.

\begin{description}
  \item[Type 1 --- Intentionally vulnerable applications.]
    Open-source projects built \emph{on purpose} with security flaws for educational or CTF use (e.g., PyGoat, DVPWA, VAmPI).
    These projects offer high vulnerability density and diverse CWE coverage, making them ideal for initial calibration.
  \item[Type 2 --- Previously-vulnerable production applications.]
    Real-world applications with known CVEs, pinned to the commit immediately \emph{before} the fix.
    Provides ecologically valid targets but requires per-CVE isolation.
  \item[Type 3 --- Previously-vulnerable libraries.]
    Library code (as opposed to applications) with CVE-linked vulnerabilities.
    Tests whether scanners can reason about library internals rather than application-level data flow.
  \item[Type 4 --- Benchmark roll-ups.]
    Existing benchmark suites (e.g., OWASP Benchmark, Juliet Test Suite) re-scored under the \textsc{RealVuln} framework to enable cross-benchmark comparison.
  \item[Type 5 --- Academic reproduction studies.]
    Datasets from published security research, re-annotated with the \textsc{RealVuln} ground truth schema for reproducibility.
\end{description}

%% ----------------------------------------------------------------
\subsection{Dataset Construction}
\label{sec:dataset-construction}

\subsubsection{Repository Selection.}
We collected 26 intentionally vulnerable Python repositories from GitHub.
Selection criteria were: (1)~the project must be a multi-file application with a meaningful code structure (single-file toy scripts were excluded), (2)~the project must use a recognizable web framework, and (3)~the project should contribute CWE families not already saturated in the dataset.
Each repository is pinned to a specific commit SHA recorded in \texttt{benchmark-manifest.json}, ensuring that all scanners analyze identical source trees.

The 26~repositories span five web frameworks: Flask (15~repos), Django (3), FastAPI (3), aiohttp (1), and Tornado (1), with 3~repos using lightweight or custom HTTP stacks.
Table~\ref{tab:dataset-stats} summarizes the dataset.

\begin{table}[t]
  \centering
  \caption{Dataset statistics for \textsc{RealVuln}~v1.0.}
  \label{tab:dataset-stats}
  \begin{tabular}{lr}
    \toprule
    \textbf{Statistic} & \textbf{Value} \\
    \midrule
    Repositories                & 26  \\
    Total labeled findings      & 796 \\
    \quad Vulnerabilities (\texttt{is\_vulnerable=true})  & 676 \\
    \quad FP traps (\texttt{is\_vulnerable=false})        & 120 \\
    Web frameworks covered      & 5   \\
    CWE families                & 18  \\
    Language                    & Python (100\%) \\
    Target type                 & Type~1 (intentionally vulnerable) \\
    \bottomrule
  \end{tabular}
\end{table}

\subsubsection{Ground Truth Labeling.}
Every finding in the ground truth was produced by manual review.
Each labeled entry records the following fields:

\begin{itemize}
  \item \textbf{Identity:} a unique \texttt{id} and a boolean \texttt{is\_vulnerable} flag.
  \item \textbf{CWE classification:} a \texttt{primary\_cwe} reflecting the most precise weakness, and an \texttt{acceptable\_cwes} list enumerating alternative CWE identifiers that a scanner may reasonably report for the same flaw (e.g., both CWE-89 and CWE-943 for a SQL injection that uses an ORM raw query).
  \item \textbf{Location:} the file path and a \texttt{start\_line}/\texttt{end\_line} range pinpointing the vulnerable code.
  \item \textbf{Severity:} one of \texttt{critical}, \texttt{high}, \texttt{medium}, or \texttt{low}.
  \item \textbf{Evidence:} the source of the annotation (\texttt{manual\_review}, \texttt{cve\_id}, or \texttt{walkthrough}) and a free-text description explaining \emph{why} the code is or is not vulnerable.
\end{itemize}

\noindent
Figure~\ref{fig:gt-entry} shows an abbreviated ground truth entry.

\begin{figure}[t]
{\footnotesize
\begin{verbatim}
{
 "id": "pygoat-001",
 "is_vulnerable": true,
 "primary_cwe": "CWE-89",
 "acceptable_cwes":
   ["CWE-89","CWE-564","CWE-943"],
 "file": "introduction/views.py",
 "location": {
   "start_line": 158,
   "end_line": 162,
   "function": "sql_lab"
 },
 "severity": "high",
 "evidence": {
   "source": "manual_review",
   "description": "SQL injection via
     string concat in raw query"
 }
}
\end{verbatim}
}
\caption{A ground truth entry for a SQL injection in PyGoat. The \texttt{acceptable\_cwes} list permits three CWE identifiers, tolerating CWE ambiguity across scanners.}
\label{fig:gt-entry}
\end{figure}

\subsubsection{Authorship Metadata.}
All 26~repositories are tagged \texttt{human\_authored} with high confidence.
These projects predate the widespread availability of large language models for code generation, which is relevant for mitigating data contamination concerns when evaluating LLM-based scanners: the ground truth code was not plausibly generated by the models under test.

\subsubsection{False Positive Traps.}
Of the 796~labeled findings, 120 (15.1\%) are \emph{FP~traps}: code patterns that appear suspicious but are demonstrably safe upon manual analysis.
For example, a login function that passes user input to SQLAlchemy's \texttt{filter\_by()} method superficially resembles SQL injection, but the ORM automatically parameterizes the query.
If a scanner flags such an entry, the match is classified as a false positive and penalized in scoring.
FP~traps serve two purposes: (1)~they test a scanner's ability to distinguish truly vulnerable code from safe-but-suspicious patterns, and (2)~they provide true negatives for computing specificity-related metrics such as the false positive rate.

%% ----------------------------------------------------------------
\subsection{Matching Algorithm}
\label{sec:matching}

Given a set of scanner findings $\mathcal{F} = \{f_1, \ldots, f_m\}$ and a ground truth set $\mathcal{G} = \{g_1, \ldots, g_n\}$, the matching algorithm determines which findings correspond to which ground truth entries.
Each finding $f$ is a triple $(\mathit{file}_f, \mathit{cwe}_f, \mathit{line}_f)$ and each ground truth entry $g$ carries $(\mathit{file}_g, \mathit{cwes}_g, [\mathit{lo}_g, \mathit{hi}_g], \mathit{vuln}_g)$, where $\mathit{cwes}_g$ is the \texttt{acceptable\_cwes} set, $[\mathit{lo}_g, \mathit{hi}_g]$ is the line range, and $\mathit{vuln}_g \in \{\mathit{true}, \mathit{false}\}$.

\paragraph{Three-field match predicate.}
A finding $f$ is a \emph{candidate match} for ground truth entry $g$ when all three conditions hold:
\begin{enumerate}
  \item \textbf{File:} $\mathit{file}_f = \mathit{file}_g$ \quad (exact match after path normalization to forward slashes).
  \item \textbf{CWE:} $\mathit{cwe}_f \in \mathit{cwes}_g$.
  \item \textbf{Line:} $\mathit{lo}_g - \delta \;\leq\; \mathit{line}_f \;\leq\; \mathit{hi}_g + \delta$, \quad where $\delta = 10$ is the default line tolerance.
        If either $\mathit{line}_f$ or $\mathit{lo}_g$ is absent, the line condition is satisfied by default (no penalty for missing location data).
\end{enumerate}

\noindent
This relaxation could advantage scanners that omit line numbers, as they effectively face a two-field match. In practice, all rule-based SAST tools provide line numbers for every finding. Among LLM scanners, 100\% of findings include line numbers --- no evaluated scanner omits location data. Kolega likewise provides line numbers for all findings. We report the fraction of line-less findings per scanner in the supplementary material.

\paragraph{Preference and consumption rules.}
When a finding matches multiple ground truth entries, the algorithm prefers entries with $\mathit{vuln}_g = \mathit{true}$, giving the scanner the benefit of the doubt (credit for detecting a real vulnerability rather than penalty for triggering a co-located trap).
Each ground truth entry is consumed at most once (greedy, first-match assignment).
Findings are processed in file-path then line-number order. While optimal (bipartite) matching would be more theoretically sound, in practice the one-to-one constraint rarely activates --- fewer than 2\% of ground-truth entries have overlapping match candidates --- making the greedy and optimal results equivalent on this dataset.

\paragraph{Classification.}
After matching, each item is classified as follows:
\begin{itemize}
  \item \textbf{TP} (true positive): finding matched a GT entry with $\mathit{vuln}_g = \mathit{true}$.
  \item \textbf{FP} (false positive): finding matched a GT entry with $\mathit{vuln}_g = \mathit{false}$, \emph{or} the finding matched no GT entry at all.
  \item \textbf{FN} (false negative): GT entry with $\mathit{vuln}_g = \mathit{true}$ that was not matched by any finding.
  \item \textbf{TN} (true negative): GT entry with $\mathit{vuln}_g = \mathit{false}$ that was not matched by any finding (scanner correctly ignored the trap).
\end{itemize}

\noindent
Algorithm~\ref{alg:matching} gives the complete procedure in pseudocode.

\begin{figure}[t]
{\footnotesize
\begin{verbatim}
MATCH(findings F, ground_truth G):
  matched_ids <- {}
  results <- []
  for each f in F:
    candidates <- []
    for each g in G:
      if g.id in matched_ids: continue
      if FILE_EQ(f, g) and CWE_IN(f, g)
         and LINE_OK(f, g):
        candidates.append(g)
    if candidates not empty:
      sort candidates: vuln=true first
      best <- candidates[0]
      matched_ids.add(best.id)
      if best.is_vulnerable:
        results.append(TP, best.id, f)
      else:
        results.append(FP, best.id, f)
    else:
      results.append(FP, null, f)
  for each g in G:
    if g.id not in matched_ids:
      if g.is_vulnerable:
        results.append(FN, g.id, null)
      else:
        results.append(TN, g.id, null)
  return results
\end{verbatim}
}
\caption{Pseudocode for the three-field matching algorithm. \texttt{FILE\_EQ} checks normalized path equality, \texttt{CWE\_IN} checks $\mathit{cwe}_f \in \mathit{cwes}_g$, and \texttt{LINE\_OK} checks line tolerance $\delta=10$.}
\label{alg:matching}
\end{figure}

%% ----------------------------------------------------------------
\subsection{Scoring}
\label{sec:scoring}

\subsubsection{Primary Metric: $F_3$ Score.}
The primary metric is the $F_3$ score, scaled to $[0, 100]$:
\begin{equation}
  F_3 = \frac{10 \cdot P \cdot R}{9P + R} \times 100
  \label{eq:f3}
\end{equation}
where $P = \frac{\mathit{TP}}{\mathit{TP} + \mathit{FP}}$ is precision and $R = \frac{\mathit{TP}}{\mathit{TP} + \mathit{FN}}$ is recall.
The $F_\beta$ family generalizes the harmonic mean with parameter $\beta$; setting $\beta = 3$ weights recall nine times more heavily than precision.

Attackers only need to exploit a single vulnerability to compromise a system. A missed SQL injection or authentication bypass can lead to complete data exfiltration, regardless of how many other issues the scanner correctly identified. This asymmetry---where a single false negative can be catastrophic while false positives merely cost analyst time---motivates an even stronger recall bias than $F_2$ provides. $F_3$ ($\beta{=}3$) weights recall $9\times$ over precision, making it the appropriate metric for production-grade security assessment. We also report $F_2$ ($\beta{=}2$, recall weighted $4\times$) for environments where noise reduction is a higher priority, such as development-time scanning or lower-risk internal tools.

We acknowledge that the choice of $\beta = 3$ favors high-recall, low-precision scanners. Under $F_1$ (equal weighting), the ranking changes substantially: scanners with higher precision would score relatively better. We report $F_2$ ($\beta = 2$) as a secondary metric throughout, and the full leaderboard data (available in the repository) allows practitioners to re-rank scanners under any $F_\beta$ of their choosing. We selected $\beta = 3$ before evaluating any scanner, based on the security domain rationale that false negatives carry asymmetrically higher cost than false positives in production environments.

\subsubsection{Secondary Metric: $F_2$ Score.}
We additionally report the $F_2$ score:
\begin{equation}
  F_2 = \frac{5 \cdot P \cdot R}{4P + R} \times 100
  \label{eq:f2}
\end{equation}
Setting $\beta = 2$ weights recall four times more heavily than precision. $F_2$ remains appropriate for lower-risk projects or development-time scanning where noise reduction matters more, but $F_3$ is the recommended default for production and high-risk environments.

\subsubsection{Additional Metrics.}
We also report precision, recall, $F_1$, the true positive rate ($\mathit{TPR} = R$), the false positive rate $\mathit{FPR} = \frac{\mathit{FP}}{\mathit{FP} + \mathit{TN}}$, and Youden's $J$ statistic ($J = \mathit{TPR} - \mathit{FPR}$).
Youden's $J$ summarizes overall discriminative ability on a $[-1, 1]$ scale, where $J = 1$ indicates perfect classification and $J \leq 0$ indicates no better than chance.

\subsubsection{Per-CWE-Family Breakdown.}
The 18~CWE families defined in \texttt{config/cwe-families.json} group related CWEs into higher-level vulnerability categories (e.g., the \emph{SQL Injection} family covers CWE-89, CWE-564, and CWE-943).
Each ground truth entry is bucketed into a family by its \texttt{primary\_cwe}, and per-family precision and recall are computed independently.
This breakdown reveals which vulnerability categories a scanner handles well and where it has blind spots.

\subsubsection{Per-Severity Breakdown.}
Similarly, TP/FP/FN counts and recall are computed per severity level (critical, high, medium, low).
This allows practitioners to assess whether a scanner prioritizes high-severity findings or exhibits uniform performance across severity tiers.

\subsubsection{Scoring Modes and Multi-Run Aggregation.}
\textsc{RealVuln} supports two aggregation modes for multi-repo evaluation:

\begin{description}
  \item[\texttt{micro}] pools TP, FP, FN, and TN counts across all repositories before computing metrics. Repositories where the scanner produced no output (e.g., due to a timeout or parsing failure) are simply omitted from the pool.
  \item[\texttt{strict\_micro}] is identical except that repositories with no scanner output are treated as if the scanner reported zero findings: every vulnerable GT entry in that repo becomes a false negative. This mode penalizes incomplete runs and is the conservative primary reporting mode in this paper. All headline scores are F3 (strict).
\end{description}

\noindent
When a scanner is run multiple times on the same repository (e.g., with different prompts or configurations), each run is scored independently.
Per-repo scores are then aggregated across runs using micro-averaging over the pooled confusion matrix.

\subsubsection{Reproducibility.}
The file \texttt{benchmark-manifest.json} locks three pieces of information necessary for exact reproduction: (1)~the content hash of the ground truth directory, (2)~the commit SHA for each of the 26~repositories, and (3)~a hash of the default prompt version used for LLM-based scanners.
Any change to the ground truth labels, repository code, or prompt template produces a new manifest, making it straightforward to detect drift between benchmark versions.
