\section{The Living Benchmark}

A static benchmark becomes stale as soon as the scanners it measures improve.
RealVuln is designed to grow alongside the tools it evaluates: new repositories,
new scanners, and new vulnerability classes can be incorporated without
invalidating prior results.

\subsection{Versioning Strategy}

Every benchmark release is identified by a \emph{manifest version}: a
deterministic hash computed from (1) the SHA-256 of every ground-truth JSON
file, (2) the upstream repository commit SHAs pinned at collection time, and
(3) the exact prompt text used for any LLM-based evaluations.  The manifest is
checked into the repository as \texttt{benchmark-manifest.json} so that any
reader can reproduce the exact evaluation context.

Scanner scores are always reported against a named manifest version (e.g.\
\texttt{v1.0.0-a3f7c2}).  When new repositories or corrected ground-truth
labels are added, the manifest hash changes and a new version tag is created.
Critically, \emph{old results remain valid}: a score computed against
\texttt{v1.0.0} continues to mean the same thing even after \texttt{v2.0.0}
is released.  The live dashboard surfaces the current version prominently and
renders side-by-side comparisons against historical versions so that score
improvements can be attributed unambiguously to scanner progress rather than
to benchmark drift.

\subsection{Roadmap}

\paragraph{Multi-language expansion.}
The \texttt{v1} corpus is Python-only.  \texttt{v2} will add intentionally
vulnerable JavaScript/TypeScript, Go, and Java repositories, enabling
language-stratified analysis and testing scanner portability.

\paragraph{Type 2 targets: production apps with CVE-linked commits.}
Rather than purpose-built educational apps, Type 2 targets are real production
applications that contained a known CVE, evaluated at the vulnerable commit.
These targets exhibit realistic vulnerability density (often a single subtle
flaw among thousands of lines of legitimate code), providing a harder and more
ecologically valid test of scanner precision.

\paragraph{Type 3 targets: vulnerable libraries.}
Many real-world security failures originate in library code rather than
application code.  Type 3 targets are open-source library releases that were
later patched for disclosed CVEs, testing whether scanners can localise
vulnerabilities in dependency-heavy, non-web codebases.

\paragraph{Type 4 roll-ups: OWASP Benchmark and Juliet.}
To enable direct comparison with existing academic benchmarks, we will
re-score OWASP Benchmark and the Juliet Test Suite using RealVuln metrics
(F2 score, per-CWE-family breakdown, FP-trap penalty).  This provides a
bridge between our recall-weighted evaluation and the precision-oriented
framing common in prior work.

\paragraph{Community scanner submissions.}
Any tool can be evaluated by dropping Semgrep-format results into the
\texttt{scan-results/} directory.  We encourage vendors and independent
researchers to submit results; all community-contributed runs are displayed on
the leaderboard with provenance metadata.

\paragraph{Agentic evaluation sandbox.}
The OpenHands Docker harness introduced in \S\ref{sec:experimental-setup}
will be extended to support fully isolated, reproducible agentic runs.
Researchers will be able to evaluate any tool-use agent against the benchmark
without requiring cloud credentials, by mounting a read-only snapshot of each
repository into a sandboxed container.

\subsection{Community Contribution}

RealVuln is maintained as a community resource.  The contribution pathways are
intentionally low-friction:

\begin{description}
  \item[Adding a repository.]
    Create a ground-truth JSON file following the schema documented in
    \texttt{ground-truth/README.md}, run \texttt{python validate\_gt.py} to
    verify schema conformance, and open a pull request.  Accepted repositories
    are assigned a version-bumping manifest update.

  \item[Adding a scanner.]
    Place Semgrep-format results in
    \texttt{scan-results/\{repo\}/\{scanner\}/results.json}.  Unknown scanner
    slugs are automatically handled by the fallback \texttt{SemgrepParser}.
    For non-Semgrep output formats, add a parser class in \texttt{parsers/}
    and register it in \texttt{PARSER\_REGISTRY}.

  \item[Challenging a ground-truth label.]
    Open a GitHub issue with evidence (e.g.\ a code path demonstrating
    exploitability, or a reference to an authoritative source).  Labels are
    reviewed by the maintainer team; accepted corrections trigger a
    patch-version manifest bump.

  \item[Leaderboard.]
    The live leaderboard is hosted at GitHub Pages and is rebuilt automatically
    on every merge to the \texttt{deploy\_pages} branch.  All scores, scanner
    metadata, and per-CWE-family breakdowns are visible without registration.
\end{description}
