\section{Conclusion}

We presented a systematic comparison of rule-based SAST tools, general-purpose
LLM scanners, and security-specialized systems, evaluated on
\textsc{RealVuln}---the first fully open-source benchmark built from real-world
vulnerable code.  Version~1 spans 26 repositories, 796 manually labeled
findings, and 120 false-positive traps across 15 scanners in three categories.
Every artifact---ground truth, raw scanner output, scoring pipeline, dashboard
generator, and evaluation prompts---is released under a permissive license
with a reproducible manifest that version-locks results to exact benchmark
snapshots.

Our experiments reveal a clear three-tier hierarchy.  The security-specialized
scanner (Kolega, F3\,=\,73.0 strict) leads the benchmark, demonstrating that
purpose-built security architecture delivers meaningfully higher recall than
general-purpose approaches.  General-purpose LLM scanners form a competitive
middle tier --- the best, Claude Sonnet~4.6 (F3\,=\,51.7), achieves nearly
3$\times$ the F3 score of the best rule-based SAST tool (Semgrep, F3\,=\,17.7).
Nevertheless, no scanner dominates across all CWE families: rule-based tools
retain a precision advantage on well-specified injection patterns, while
LLM-based systems excel on logic-level and authentication flaws that require
broader code context.  The contrast between the two security-specialized systems reinforces this conclusion: the SecLab Taskflow Agent, despite sophisticated multi-stage threat modeling with GPT-5.4, scores F3\,=\,7.1---below every rule-based SAST tool---because its depth-first design audits only 3--5 threat categories per repository. Purpose-built breadth of coverage, not merely security-specialized architecture, is what separates top-tier detection from the rest.
These results suggest that the future of vulnerability
detection lies not in applying general-purpose AI to security prompts, but in
building security-specialized systems that combine LLM reasoning with
domain-specific architecture \emph{and} broad detection coverage.

The benchmark is designed to evolve.  A community contribution pathway,
described in \S\ref{sec:living-benchmark}, allows new repositories, scanners,
and ground-truth corrections to be incorporated without invalidating prior
results.  Version 2 will expand coverage to JavaScript/TypeScript, Go, and
Java, and will introduce Type 2 targets drawn from real production CVEs ---
setting a higher bar for precision in realistic, low-vulnerability-density
codebases.  We invite the security research community to submit scanner results,
propose new repositories, and challenge existing labels.

All code, data, results, and tooling are available at \url{https://github.com/kolega-ai/RealVulnBenchmark}. A live interactive dashboard is hosted at \url{https://realvuln.kolega.dev}. We encourage the security community to verify our results, challenge our ground truth, and submit their own scanner evaluations.
